---
title: "Classification Template"
author: "@eisamahyari"
date: "`r Sys.Date()`"
output:
  rmdformats::html_clean:
    fig_width: 14
    fig_height: 10
    highlight: kate
    thumbnails: false
    lightbox: false
    gallery: false
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

# Introduction

This is a template to use RIRA for classification of cell types 

# libs


```{r setup, include=FALSE}
DevMode = T


library(knitr)
library(rmdformats)

## Global options
options(max.print="75")

opts_chunk$set(echo=T,
	             cache=F,
               prompt=F,
               tidy=T,
               comment=NA,
               message=F,
               warning=F)

opts_knit$set(width=75)
knitr::opts_chunk$set(fig.width=9, fig.height=6)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```


```{r}

library(Seurat)

library(ggplot2)
library(dplyr)
library(Matrix)
library(viridis)
library(DT)
library(patchwork)

library("BiocParallel")
register(MulticoreParam(6))


library(CellMembrane)
library(scCustFx)
library(RIRA)

col_vector <- scCustFx::ColorTheme()$col_vector
ggtheme = scCustFx:::theme_simple
```
``

# data input

## RIRA

Load in the entire RIRA and add precomputed Kfolds meta then split and save the needed k-fold. 

piloting on RIRA_f1_10 which is fold 1 of 10 .

once saved its not necessary to load the giant full RIRA. 

```{r}
# RIRA <- readRDS("../RhesusAtlas/data/serobjs/primeV2/SingleCell.merge.ds.norm.pca.dr.markers.singler.dimplot.phenotype.cite-pca.wnn.featureplot.cite-plots.RhesusAtlas2_7TissueComboV1.seurat.rds")
# 
# RIRA$barcode = colnames(RIRA)
# 
# Tissue_KfoldLS = readRDS("../RhesusAtlas/data/RIRA_full_KfoldLS.rds")
# 
# RIRA$K5fold = 0 
# for (ifold in 1:length(Tissue_KfoldLS$Tissue_Kfold_LS5)){
#   print(ifold)
#   RIRA$K5fold[Tissue_KfoldLS$Tissue_Kfold_LS5[[ifold]]] =paste0("Kfold",ifold)
# }
# RIRA$K10fold = 0 
# for (ifold in 1:length(Tissue_KfoldLS$Tissue_Kfold_LS10)){
#   print(ifold)
#   RIRA$K10fold[Tissue_KfoldLS$Tissue_Kfold_LS10[[ifold]]] =paste0("Kfold",ifold)
# }
# table(RIRA$K10fold)
# 
# RIRA_f1_10 = subset(RIRA, K10fold == "Kfold1")
# dim(RIRA)
# dim(RIRA_f1_10)
# DimPlot(RIRA_f1_10)
# slot(object = RIRA_f1_10[["RNA"]], name = "scale.data") <- new(Class = "matrix")
# slot(object = RIRA_f1_10[["ADT"]], name = "scale.data") <- new(Class = "matrix")

# library(SeuratDisk)
# SaveH5Seurat(RIRA_f1_10, "../RhesusAtlas/data/serobjs/RIRA_f1_10", overwrite = F)

# saveRDS(RIRA_f1_10, "../RhesusAtlas/data/serobjs/RIRA_f1_10.rds")
# remove(RIRA)
# remove(Tissue_KfoldLS)
```

### subsets

piloting on RIRA_f1_10 which is fold 1 of 10 . Custom fx DietSeurat_Cust reduces the object to the needed sub objects. Training_epochs are neeeded sometimes, here we randomize it and keeping the sizes balanced, but it can be based on any metadata. 

```{r}
# RIRA_f1_10b = LoadH5Seurat("../RhesusAtlas/data/serobjs/RIRA_f1_10.rds.h5seurat")
RIRA_f1_10 = readRDS("../RhesusAtlas/data/serobjs/RIRA_f1_10.rds")
# RIRA_f1_10[["ADT"]] = NULL
DimPlot(RIRA_f1_10)


set.seed(1234)

RIRA_f1_10 = DietSeurat_Cust(RIRA_f1_10, addCITE = F)

RIRA_f1_10$Training_epochs = factor(sample(LETTERS[1:5], length(RIRA_f1_10$barcode), replace = T))

```

## Query sets

### SIV pilot 1

load in a new data of interest

```{r}

Query1 =readRDS("../SIVacute/data/serobjs/prime/SingleCell.ds.merge.norm.pca.dr.markers.singler.Pilot_Timpoint_TresRh_DS5000.seurat.rds")

Query1$barcode = colnames(Query1)

Query1 = DietSeurat_Cust(Query1, addCITE = F)

Query1$Testing_epochs = factor(sample(LETTERS[1:10], length(Query1$barcode), replace = T))

table(Query1$Testing_epochs )



```



# Classification / label transfer

# I/O for classifiers

```{r}
RefLS =  SplitObject(RIRA_f1_10, split.by = "Training_epochs")

Query1_f1of5 = subset(Query1, Testing_epochs %in% c("A"))

Query1_f1of5 = Query1_f1of5 #%>%  NormalizeData(verbose = T) %>% FindVariableFeatures(selection.method = "vst", nfeatures = 2000, verbose = T)


# Query1_f1of5 = Rdiscvr::DownloadAndAppendTcrClonotypes(Query1_f1of5)
# Query1_f1of5 = Rdiscvr::QueryAndApplyCdnaMetadata(Query1_f1of5)
# Query1_f1of5 = scCustFx::ReduceSerObj_HTO(Query1_f1of5)


```



## Seurat CCA

### Training of Ref

```{r}
table(RIRA_f1_10$Training_epochs)
if(!file.exists("./data/Ref_IntLS_v1.rds")){
  IntLs = integrate_reflist_CCA(RefLS = RefLS, dims = 1:30, dims_umap=1:20)
  saveRDS(IntLs, "./data/Ref_IntLS_v1.rds")
} else {
  IntLs = readRDS("./data/Ref_IntLS_v1.rds")
}

Ref.anchors = IntLs$Ref.anchors
Ref.integrated = IntLs$Ref.integrated

remove(IntLs)
```


### Mapping of Query

```{r}
QueryLs = classifier_Seurat(Ref.integrated = Ref.integrated, Query.Ser=Query1_f1of5, dims = 1:30, dims_umap=1:20,
                              Ref.meta= Ref.integrated$Pheno.scope1)
  
Query1_f1of5 = QueryLs$Query.Ser
Query_anchors = QueryLs$Query_anchors
remove(QueryLs)

#extra meta feats

predictions <- TransferData(anchorset = Query_anchors,
                            refdata = Ref.integrated$Pheno.scope2,
                            dims = 1:20)

# head(predictions)
# pheatmap::pheatmap(t(predictions[sample(1:nrow(predictions), 100),2:(ncol(predictions)-1)]))

Query1_f1of5 <- AddMetaData(Query1_f1of5, metadata = predictions, col.name = "Pheno.scope2")

```

## cFIT

```{r}
# devtools::install_github("pengminshi/cFIT")
library(cFIT)

RefLS$A@meta.data$B_ID = sample(LETTERS[1:5], nrow(RefLS$A@meta.data), replace = T)
table(RefLS$A@meta.data$B_ID )

data.list <- format_refdata(ref_seurat = RefLS$A, ref_meta_batchID = RefLS$A$B_ID, 
                            ref_meta = RefLS$A@meta.data, 
                            ref_meta_labels = RefLS$A@meta.data[,c("dice.label")], 
                            dataset.name = 'RIRA:')

# data.list <- format_refdata(RIRA_f1_10, RIRA_f1_10$Training_epochs, RIRA_f1_10@meta.data)

# select highly variable genes in reference
cFIT_topVargenes = cFIT::select_genes(data.list$X.list, ngenes=500, verbose=F)

if(!file.exists("./data/Ref_cFITv1.rds")){
  int.out = integrate_ref(genes=cFIT_topVargenes, data.list=data.list)
  saveRDS(int.out, "./data/Ref_cFITv1.rds")
} else {
  int.out = readRDS("./data/Ref_cFITv1.rds")
}

if(!file.exists("./data/Query1_f1of5_cFITv1.rds")){
  tf.out = transfer_sigs(genes=cFIT_topVargenes, target_seurat=Query1_f1of5, int.out=int.out) 
  saveRDS(tf.out, "./data/Query1_f1of5_cFITv1.rds")
} else {
  tf.out = readRDS("./data/Query1_f1of5_cFITv1.rds")
}




#assign labels from reference to the target data
est.labels = cFIT::asign_labels(exprs.source=do.call(rbind, int.out$H.list),
                                exprs.target=tf.out$H,
                                labels.source=do.call(c, data.list$labels.list))


Query1_f1of5$predicted_lab = est.labels



Query1_f1of5@meta.data$rnaTSNE_1 =  Query1$rnaTSNE_1[rownames(Query1_f1of5@meta.data)]
Query1_f1of5@meta.data$rnaTSNE_2 =  Query1$rnaTSNE_2[rownames(Query1_f1of5@meta.data)]
Query1$predicted_lab = "unk"
Query1$predicted_lab[rownames(Query1_f1of5@meta.data)] =  Query1_f1of5$predicted_lab

FeatureScatter(Query1_f1of5, feature1 = "rnaTSNE_1", feature2 = "rnaTSNE_2", group.by = "predicted_lab")
FeatureScatter(Query1, feature1 = "rnaTSNE_1", feature2 = "rnaTSNE_2", group.by = "predicted_lab", cols = col_vector)

```

## SignleR

```{r}

```

